---
title: "Dataset"
output: html_document
---

```{r}
#Generating censoring time using Kaplan Meier estimator with uniform noise.

CenTime <- function(X, Status, M){
  #X: observed survival time
  #Status: uncensored=1, censored=0
  
  n <- length(X)
  res <- X
  obs <- seq(from = 1, to  = n)[Status == 1]
  Status0 <- Status
  X0 <- X
  
  #sort the data by observed survival time
  Status <- Status[order(X)]
  X <- sort(X)
  
  ind <- 1:n
  LHS <- ((n-ind)/(n-ind+1))^Status
  LHS <- cumprod(LHS)
  KMEst <- LHS[Status == 0]#LHS
  
  Nc <- n - sum(Status)#number of censored cases
  
  Tcen <- X[Status == 0]#vector of censoring time
  num_largeu <- 0#record the time when the inequality doesn't hold
  
  
  for (i in obs) {
    Ci <- min(seq(Nc)[KMEst <= runif(1)])
    if(Ci == Inf){
        res[i] <- Tcen[Nc]#in case min(KMEst) > u[i]
        num_largeu <- num_largeu + 1
      }
    else{
        res[i] <- Tcen[Ci]
      }
      
    #conditiong on generated censoring time is larger than the survival time.
    if (res[i] <= X0[i]){
        res[i] <- runif(n = 1, min = X0[i], max = M)
        #the observed survival time must be smaller than the upperbound of censoring distribution
      }
    
  }
  
  return(list("res" = jitter(res, factor = 2),
              "urate" = num_largeu/n))
}
```

```{r}
#Generating survival time by Breslow-type estimator 
#using beta_hat, covariate, observed survival time and Status.
#the output is ordered by corresponding covariate.

SurvTime <- function(beta, MySimu){
  #Z: covariate
  #X: observed survival time.
  #Status: uncensored=1, censored=0.
  #Z = MySimu_ori$covs, X= MySimu_ori$stop, Status = MySimu_ori$event
  
  #sort by observed survival time to evaluate the integral
  MySimu <- MySimu[order(MySimu$stop),]
  Z <- MySimu$covs
  Status <- MySimu$event
  X <- MySimu$stop
  num_largeu <- 0
  
  samp_size <- length(Z)
  nstar <- sum(Status)
  Tstar <- X[Status == 1]
 
  Baseline <- -log(runif(samp_size))/exp(Z*beta)
  
  risk <- cumsum(Status)
  denom <- c()
  for (i in 1:nstar) {
    denom[i] <- sum(exp(beta*Z)[risk >= i])
  }
  BSEst <- cumsum(1/denom)
  
  res <- c()
  for (i in 1:samp_size) {
    Ti <- min(seq(nstar)[BSEst >=Baseline[i]])
    if(Ti == Inf){
      res[i] <- Tstar[nstar]##in case max(BSEst)<Baseline[i]
      num_largeu <- num_largeu + 1
    }
    else{
      res[i] <- Tstar[Ti]
    }
  }
  
  MySimu$X <- res
  MySimu <- MySimu[order(MySimu$covs),]
  
  #smoothing survival time
  observed <- seq(samp_size)[MySimu$event > 0]
  for (i in 1:(nstar-1)) {
    nums <- observed[i] - observed[i+1]
    if(nums >= 2){
      MySimu$X[(observed[i]+1) : (observed[i+1]-1)] <- sort(runif(nums - 1, 
                                                       min = MySimu$X[observed[i]], 
                                                       max = MySimu$X[observed[i+1]]), decreasing = TRUE)
    }
  }
  MySimu$X[observed[i+1]:samp_size] <- jitter(MySimu$X[observed[i+1]:samp_size], factor = 2)
  return(list("res" = MySimu$X,
              "urates" = num_largeu/samp_size))
}

```


```{r}
############Not work for failures at the same time
nlpl_Censor_ind <- function(events, surv_time, uni_cov, zk, mono_f = NA, est_beta = NA){
  # flag = get("flag", envir = .GlobalEnv)
  # if(flag){
  #   return (1)
  # }
  #events : 0,1 indicating whether an observed failure happened (vector)
  #surv_time : observed survival time (vector)
  #uni_cov : covariate (vector)
  #mono_f : estimated phi value using isoph in isoSurv package (vector). It is sorted by covariate when output by isoph!
  #est_beta : estimated parameter for linear hazard if applicable (scalar)

  z <- uni_cov[order(surv_time)]
  Status <- events[order(surv_time)]
  mono_f <- mono_f[order(surv_time)]
  X <- sort(surv_time)
  star <- Status
  n_star <- sum(star)

  #partial likelihood
  denominator <- c()
  risk_set <- cumsum(star)
  #monotone
  if(is.na(est_beta)){
    ll1 <- -sum(mono_f*star)
    for(i in 1:n_star){
      denominator[i] <- (risk_set == i)%*%exp(mono_f)
    }
    ll2 <- sum(log(rev(cumsum(rev(denominator)))))
  }
  
  #linear
  if(is.na(mono_f)[1]){
    #denominator
    ll1 <- -sum(est_beta*(z - zk)*star)
    for(i in 1:n_star){
      denominator[i] = (risk_set >= i)%*%exp(est_beta*(z - zk))#sum for all T_j > = T_i(exp(beta*z_j))
    }
    ll2 <- sum(log(denominator))
  }
  return(ll1+ll2)
}

```











# 1

```{r}
library(BSDA)
summary(Aids)
#plot(x = Aids$age, y = Aids$duration)
```
Incubation times for 295 patients thought to be infected with HIV by a blood transfusion.
Source: Kalbsleich, J. and Lawless, J., (1989), An analysis of the data on transfusion related AIDS.


```{r, warning=FALSE}
# source("Bootstrap_c_exp.R")
# source("Bootstrap_c_gp.R")
# source("Bootstrap_c_wb.R")
#source("Bootstrap_exp_SSurvFixKM_moutofn.R")
# source("nlpl_noCensor_ind.R")
# source("SurvTimeSmoothSurv.R")
# source("CenTimeFix.R")



B_rep = 1  #500
B_size = 200

library(survival)
library(isoSurv)
library(pracma)

source('myisoph_initial.R')
assignInNamespace("isoph.initial", isoph.initial, pos = "package:isoSurv")

library(BSDA)
mydata <- Aids
sampsize_ori = dim(mydata)[1] #500



#Generate the original sample
sig_level <- 0.05
tf <- FALSE #timefix

MySimu_ori <- data.frame("stop" = mydata$duration, "covs" = mydata$age, "event" = rep(1,sampsize_ori))

##Start from the first(in covariate) observed failure
N <- length(mydata$age)
MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
location_start <- min(seq(N)[MySimu_ori$event > 0])
MySimu <- MySimu_ori[location_start:N,]

#Estimate beta and phi
iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
psi_hat <- iso$iso.cov$psi.hat
nlpl_iso0 <- nlpl_noCensor_ind(
                             surv_time = MySimu$stop, 
                             mono_f = psi_hat, 
                             uni_cov = MySimu$covs, 
                             zk = iso$Zk)

cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
mybeta <- max(cox$coefficients,0)
nlpl_cox0 <- nlpl_noCensor_ind(
                             surv_time = MySimu$stop, 
                             est_beta = mybeta, 
                             uni_cov = MySimu$covs,
                             zk = iso$Zk)

Tobs <- nlpl_iso0 - nlpl_cox0


#Generate bootstrap sample
Tn <- c()

for (i in 1:B_size) {
  covs_B <- MySimu_ori$covs
  hzrate_B <- mybeta*covs_B
  
  survgen <- SurvTime(beta = mybeta, MySimu_ori)
  survtime_B <- survgen$res
  
  #cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
  #centime_B <-cengen$res
  
  Status_B <- rep(1, length(covs_B))
  X_B <- survtime_B
  
  MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
  MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
  location_start <- min(seq(N)[MySimu_B$event > 0])
  MySimu_B <- MySimu_B[location_start:N,]
  
  isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
  nlpl_iso <- nlpl_noCensor_ind(
                              mono_f = isofit$iso.cov$psi.hat, 
                              uni_cov = MySimu_B$covs, 
                              surv_time = MySimu_B$stop, 
                              zk = isofit$Zk)
  
  coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
  beta_hat <- max(0, coxfit$coefficients)
  nlpl_cox <- nlpl_noCensor_ind(
                              est_beta = beta_hat, 
                              uni_cov = MySimu_B$covs, 
                              surv_time = MySimu_B$stop, 
                              zk = isofit$Zk)
  
  Tn[i] <- nlpl_iso - nlpl_cox
}

T_critical <- quantile(Tn, sig_level)


results200 <- list(
  "T_obs" = Tobs,
  "T_n" = Tn,
  "beta_hat" = mybeta,
  "T_critical" = T_critical,
  "psi_hat" = iso$iso.cov$psi.hat,
  "Z" = iso$iso.cov$covs,
  "p_value" = mean(as.numeric(Tn < Tobs)))


save(results200, file = "BSDA_Aids_SSurvFixKM200.rda")

T_critical

Tobs

beta_hat

```




# 2
"cal8082.dat" in SAS. (Haven't find the data yet.)


# 3
Heart failure data: 

https://archive.ics.uci.edu/ml/datasets/Heart+failure+clinical+records

related papers: 

https://bmcmedinformdecismak.biomedcentral.com/track/pdf/10.1186/s12911-020-1023-5.pdf

https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0181001
```{r}
library(survival)
library(isoSurv)

HFData <- na.omit(read.csv("D:/0My research/data/heart_failure_clinical_records_dataset.csv"))
#library(isoph)
fit1 <- coxph(Surv(time, DEATH_EVENT)~., HFData)
summary(fit1)
```

censoring rate

```{r}
1-mean(HFData$DEATH_EVENT)
```

```{r}
cor(HFData)
```


```{r}
fit2 <- coxph(Surv(time, DEATH_EVENT)~age, HFData)
summary(fit2)
fit3 <- coxph(Surv(time, DEATH_EVENT)~serum_creatinine, HFData)
summary(fit3)
fit4 <- coxph(Surv(time, DEATH_EVENT)~ejection_fraction, HFData)
summary(fit4)
```


Testing linearity of serum creatinine.

```{r, warning = FALSE}
library(isoSurv)
B_rep = 1  #500
B_size = 500
                   
sampsize_ori = dim(HFData)[1] 
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = HFData$time, "covs" = HFData$serum_creatinine, "event" = HFData$DEATH_EVENT)

  ##Start from the first(in covariate) observed failure
  N <- sampsize_ori
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  plot(x = MySimu$covs - iso$Zk, y = psi_hat, type = 'l')
  lines(x = MySimu$covs - iso$Zk, y = mybeta*(MySimu$covs - iso$Zk))
  print(iso$Zk)
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  print("critical value =", T_critical)
  
  print("Observed value for test statistic=", Tobs)
  
  print("beta_hat = ", beta_hat)
  
  print("p_value=", mean(as.numeric(Tn < Tobs)))
 
#save(results200, file = "dataSSurvFixKM200.rda")

```
Plot comparing linear estimator and monotone estimator.

```{r}
plot(x = iso$iso.cov$covs - iso$Zk, y = psi_hat, xlab = 'Z-Zk', ylab = 'Fitted value', col = "red", type = "b", pch = 20, lty = 2)
lines(x = iso$iso.cov$covs - iso$Zk, y = mybeta*(iso$iso.cov$covs - iso$Zk), col = "blue", type = "l", lty = 1, pch = 20)
legend("topleft", legend = c("monotone hazard model", "Cox model"), lty = c(2,1), col = c("red", "blue"), cex = 1.2)
```

```{r}
##Partial Linear
coxfit <- coxph(Surv(time, DEATH_EVENT) ~ ., HFData)
isofit <- isoph(Surv(time, DEATH_EVENT) ~ iso(serum_creatinine, shape = "inc") 
                + age + anaemia + creatinine_phosphokinase + diabetes + ejection_fraction + high_blood_pressure
                + platelets + serum_sodium + sex+smoking, data = HFData)

beta_hat <- coxfit$coefficients[8]#betahat for serum creatinine
plot(x = isofit$iso.cov$serum_creatinine - isofit$Zk, 
     y = isofit$iso.cov$psi.hat, 
     xlab = 'Z-Zk', ylab = 'Fitted value', 
     col = "red", type = "b", pch = 20, lty = 2)
lines(x = isofit$iso.cov$serum_creatinine - isofit$Zk, 
      y = beta_hat*(isofit$iso.cov$serum_creatinine - isofit$Zk), 
      col = "blue", type = "l", lty = 1, pch = 20)
legend("topleft", legend = c("monotone hazard model", "Cox model"), lty = c(2,1), col = c("red", "blue"), cex = 1.2)
```


Testing linearity of ejection fraction.

```{r, warning=FALSE}

B_rep = 1  #500
B_size = 200
 
                   
mydata <- na.omit(HFData)
sampsize_ori = dim(mydata)[1] 
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = mydata$time, "covs" = -mydata$ejection_fraction, "event" = mydata$DEATH_EVENT)

  ##Start from the first(in covariate) observed failure
  N <- sampsize_ori
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  T_critical
  
  Tobs
  
  beta_hat


```


Testing linearity of age.

```{r, warning=FALSE}

B_rep = 1  #500
B_size = 200
 
                   
mydata <- na.omit(HFData)
sampsize_ori = dim(mydata)[1] 
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = mydata$time, "covs" = mydata$age, "event" = mydata$DEATH_EVENT)

  ##Start from the first(in covariate) observed failure
  N <- sampsize_ori
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  T_critical
  
  Tobs
  
  beta_hat


```

# 4. 

lung cancer dataset in 1994.
Survival in patients with advanced lung cancer from the North Central Cancer Treatment Group.
Performance scores rate how well the patient can perform usual daily activities.
nst: Institution code
time: Survival time in days
status: censoring status 1=censored, 2=dead
age: Age in years
sex: Male=1 Female=2
ph.ecog: ECOG performance score as rated by the physician. 0=asymptomatic, 1= symptomatic but completely ambulatory, 2= in bed <50% of the day, 3= in bed > 50% of the day but not bedbound, 4 = bedbound
ph.karno: Karnofsky performance score (bad=0-good=100) rated by physician
pat.karno: Karnofsky performance score as rated by patient
meal.cal: Calories consumed at meals
wt.loss: Weight loss in last six months

```{r}
data(cancer, package="survival")
HCData <- cancer
cancer$status <- cancer$status - 1#dead = 1, censored = 0
fit_lung <- coxph(Surv(time, status)~., HCData)
summary(fit_lung)
plot(cancer$age, cancer$time)
plot(cancer$ph.ecog, cancer$time)
```



```{r}
fit_lung1 <- coxph(Surv(time, status)~age, HCData)
summary(fit_lung1)
```

```{r}
fit_lung2 <- coxph(Surv(time, status)~ph.ecog, HCData)
summary(fit_lung2)
```

```{r, warning=FALSE}
# source("Bootstrap_c_exp.R")
# source("Bootstrap_c_gp.R")
# source("Bootstrap_c_wb.R")
#source("Bootstrap_exp_SSurvFixKM_moutofn.R")

#source("nlpl_Censor_ind.R")
#source("SurvTimeSmoothSurv.R")
#source("CenTimeFix.R")



B_rep = 1  #500
B_size = 200

# library(survival)
# library(isoSurv)
# library(pracma)

# source('myisoph_initial.R')
# assignInNamespace("isoph.initial", isoph.initial, pos = "package:isoSurv")
                   
mydata <- na.omit(HCData)
sampsize_ori = dim(mydata)[1] #500
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = mydata$time, "covs" = mydata$ph.ecog, "event" = mydata$status-1)

  ##Start from the first(in covariate) observed failure
  N <- length(mydata$ph.ecog)
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  T_critical
  
  Tobs
  
  beta_hat
  
 
#save(results200, file = "dataSSurvFixKM200.rda")

```











# 5. 
Monoclonal gammopathy data in 1993 and updated to correct small errors in 2015.
Natural history of 241 subjects with monoclonal gammopathy of undetermined significance (MGUS).
id: subject id
age: age in years at the detection of MGUS
sex: male or female
dxyr: year of diagnosis
pcdx: for subjects who progress to a plasma cell malignancy
the subtype of malignancy: multiple myeloma (MM) is the
most common, followed by amyloidosis (AM), macroglobulinemia (MA),
and other lymphprolifative disorders (LP)
pctime: days from MGUS until diagnosis of a plasma cell malignancy
futime: days from diagnosis to last follow-up
death: 1= follow-up is until death
alb: albumin level at MGUS diagnosis
creat: creatinine at MGUS diagnosis
hgb: hemoglobin at MGUS diagnosis
mspike: size of the monoclonal protein spike at diagnosis

```{r}
library(survival)
data0 <- mgus

```
```{r}
data1 <- mgus1
```

```{r}
# mgus#time independent
# mgus1#time dependent

MGData <- na.omit(mgus)
fit_mg <- coxph(Surv(futime, death)~., MGData)
summary(fit_mg)

MGData1 <- na.omit(mgus1)
fit_mg1 <- coxph(Surv(start, stop, status)~., MGData1)
summary(fit_mg1)

fit_mg2 <- coxph(Surv(futime, death)~pctime, MGData)#sig
summary(fit_mg2)

fit_mg3 <- coxph(Surv(futime, death)~mspike, MGData)#not sig
summary(fit_mg3)

fit_mg4 <- coxph(Surv(futime, death)~age, MGData)#sig
summary(fit_mg4)
```


Testing linearity of pctime.

```{r, warning=FALSE}
# source("Bootstrap_c_exp.R")
# source("Bootstrap_c_gp.R")
# source("Bootstrap_c_wb.R")
#source("Bootstrap_exp_SSurvFixKM_moutofn.R")

#source("nlpl_Censor_ind.R")
#source("SurvTimeSmoothSurv.R")
#source("CenTimeFix.R")



B_rep = 1  #500
B_size = 200

# library(survival)
# library(isoSurv)
# library(pracma)

# source('myisoph_initial.R')
# assignInNamespace("isoph.initial", isoph.initial, pos = "package:isoSurv")
                   
sampsize_ori = dim(MGData)[1] #500
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = MGData$futime, "covs" = MGData$pctime, "event" = MGData$death)

  ##Start from the first(in covariate) observed failure
  N <- sampsize_ori
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  T_critical
  
  Tobs
  
  beta_hat
  
 
#save(results200, file = "dataSSurvFixKM200.rda")

```

Testing linearity of age.

```{r, warning=FALSE}
# source("Bootstrap_c_exp.R")
# source("Bootstrap_c_gp.R")
# source("Bootstrap_c_wb.R")
#source("Bootstrap_exp_SSurvFixKM_moutofn.R")

#source("nlpl_Censor_ind.R")
#source("SurvTimeSmoothSurv.R")
#source("CenTimeFix.R")



B_rep = 1  #500
B_size = 200

# library(survival)
# library(isoSurv)
# library(pracma)

# source('myisoph_initial.R')
# assignInNamespace("isoph.initial", isoph.initial, pos = "package:isoSurv")
                   
sampsize_ori = dim(MGData)[1] #500
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = MGData$futime, "covs" = MGData$age, "event" = MGData$death)

  ##Start from the first(in covariate) observed failure
  N <- sampsize_ori
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  T_critical
  
  Tobs
  
  beta_hat
  
 
#save(results200, file = "dataSSurvFixKM200.rda")

```


# 6.


```{r}
library(JM)
# mydata_tind <- aids.id
# mydata_td <- aids
```

```{r}
fit_aids <- coxph(Surv(stop, event)~., aids.id[-1])
summary(fit_aids)

fit_aids1 <- coxph(Surv(stop, event)~CD4, aids.id)
summary(fit_aids1)
```

Testing linearity of CD4 cell count.

```{r, warning = FALSE}
B_rep = 1  #500
B_size = 500
                   
sampsize_ori = dim(aids.id)[1] 
                    

  
  #Generate the original sample
  sig_level <- 0.05
  tf <- FALSE #timefix
  
  MySimu_ori <- data.frame("stop" = aids.id$Time, "covs" = -aids.id$CD4, "event" = aids.id$death)

  ##Start from the first(in covariate) observed failure
  N <- sampsize_ori
  MySimu_ori <- MySimu_ori[order(MySimu_ori$covs),]
  location_start <- min(seq(N)[MySimu_ori$event > 0])
  MySimu <- MySimu_ori[location_start:N,]
  
  #Estimate beta and phi
  iso <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu)
  psi_hat <- iso$iso.cov$psi.hat
  nlpl_iso0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               mono_f = psi_hat, 
                               uni_cov = MySimu$covs, 
                               zk = iso$Zk)
  
  cox <- coxph(Surv(stop, event) ~ covs, data = MySimu, control = coxph.control(timefix = tf), ties = "breslow")
  mybeta <- max(cox$coefficients,0)
  nlpl_cox0 <- nlpl_Censor_ind(events = MySimu$event, 
                               surv_time = MySimu$stop, 
                               est_beta = mybeta, 
                               uni_cov = MySimu$covs,
                               zk = iso$Zk)
  
  Tobs <- nlpl_iso0 - nlpl_cox0
  
  plot(x = MySimu$covs - iso$Zk, y = psi_hat, type = 'l')
  lines(x = MySimu$covs - iso$Zk, y = mybeta*(MySimu$covs - iso$Zk))
  
  
  #Generate bootstrap sample
  Tn <- c()
  
  for (i in 1:B_size) {
    covs_B <- MySimu_ori$covs
    hzrate_B <- mybeta*covs_B
    
    survgen <- SurvTime(beta = mybeta, MySimu_ori)
    survtime_B <- survgen$res
    
    cengen <- CenTime(X= MySimu_ori$stop, Status = MySimu_ori$event, M=max(MySimu_ori$stop))
    centime_B <-cengen$res
    
    Status_B <- as.numeric(survtime_B <= centime_B)
    X_B <- pmin(survtime_B, centime_B)
    
    MySimu_B <- data.frame("stop" = X_B, "covs" = covs_B, "event" = Status_B)
    MySimu_B <- MySimu_B[order(MySimu_B$covs),]#sort by covariate to drop the records with covariate smaller than the covariate of the first observed failure
    location_start <- min(seq(N)[MySimu_B$event > 0])
    MySimu_B <- MySimu_B[location_start:N,]
    
    isofit <- isoph(Surv(stop, event) ~ iso(covs, shape = "inc"), data = MySimu_B)
    nlpl_iso <- nlpl_Censor_ind(events = MySimu_B$event,
                                mono_f = isofit$iso.cov$psi.hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    coxfit <- coxph(Surv(stop, event) ~ covs, data = MySimu_B, control = coxph.control(timefix = tf), ties = "breslow")
    beta_hat <- max(0, coxfit$coefficients)
    nlpl_cox <- nlpl_Censor_ind(events = MySimu_B$event,
                                est_beta = beta_hat, 
                                uni_cov = MySimu_B$covs, 
                                surv_time = MySimu_B$stop, 
                                zk = isofit$Zk)
    
    Tn[i] <- nlpl_iso - nlpl_cox
  }
  
  T_critical <- quantile(Tn, sig_level)
  
  
  T_critical
  
  Tobs
  
  beta_hat
  
 
#save(results200, file = "dataSSurvFixKM200.rda")

```

```{r}
plot(x = MySimu$covs - iso$Zk, y = psi_hat, xlab = 'Z-Zk', ylab = 'Fitted value', col = "red", type = "b", pch = 20, lty = 2)
lines(x = MySimu$covs - iso$Zk, y = mybeta*(MySimu$covs - iso$Zk), col = "blue", type = "l", lty = 1, pch = 20)
legend("topleft", legend = c("monotone hazard model", "Cox model"), lty = c(2,1), col = c("red", "blue"), cex = 1.2)
```

```{r}
#in JM package
#pbc2.id is time independent
#pbc2 is time dependent

```